\subsection{Governing equations}
\label{sec:governing-equations}

%% \paragraph{Ideal Magneto-Hydrodynamic Equations}
%% The magneto-hydrodynamic equations (MHD) are a corner stone of theoretical
%% astrophysics.  They model the fluid mechanics of ionized interstellar media in
%% an idealized form. In CGS unit they read as follows.
%% 
%% % \begin{align}
%% % \partial_t\dens + \nabla \cdot (\dens\,\velv)   &=  0 \\
%% % \dens\, \partial_t \velv + \dens(\velv \cdot \nabla) \velv + \nabla \pres &= \frac{\curdensv \times \magdensv}{c} + \dens \vec{g} + \vec{F} \\
%% % \nabla \times \magdensv &= \frac{4\;\pi}{c} \curdensv \\
%% % \nabla \times \elefield &= -\frac{1}{c} \partial_t \magdensv \\
%% % \nabla \cdot  \magdensv &= 0 \\
%% % \curdensv               &= \sigma \left ( \elefield + \frac{\velv \times \magdensv}{c} \right )
%% % \end{align}
%% 
%% \begin{equation*}
%%     \frac{1}{\partial t} \left[ \begin{array}{c}
%%         \rho \\
%%         \rho\,\vec{v} \\
%%         E \\
%%         \vec{B}
%%     \end{array} \right] +
%%     \nabla \cdot \left[ \begin{array}{c}
%%         \rho\,\vec{v}\\
%%         \rho\,(\vec{v}\otimes\vec{v}) + (p + \frac{1}{2} ||\vec{B}||^2)I - \vec{B} \otimes \vec{B} \\
%%         \vec{v}(E + p + \frac{1}{2} ||\vec{B}||^2) - \vec{B} (\vec{v}\cdot\vec{B}) \\
%%         \vec{v} \otimes \vec{B} - \vec{B} \otimes \vec{v}
%%     \end{array} \right] + \text{forcing} = 0, \;\; \nabla \cdot \vec{B} = 0,
%% \end{equation*}
%% where $\mat{a} \otimes \mat{b}$ is the \vip{Kronecker} product between two
%% matrices $\mat{a}$ and $\mat{b}$. For vectors it can also be written as
%% $\vec{a}\,\vec{b}^T$.
%% 
%% TODO: Explain above equations in detail. Name -> physical meaning ...
%% 
%% We want to keep the simulation simple and do not include electric current
%% ($\sigma = \infty$) and gravity in our model. Hence, $\curdensv = 0$ and
%% $\vec{g} = 0$. This simplifies above equations to the ideal MHD equations.
%% They are valid for dynamics in interstellar clouds where we have huge
%% spatial dimensions of several parsecs and consequently long crossing times.
%% 
%% TODO: More arguments why ideal MHD equations are valid for ISM simulations
%% 
%% TODO: Insert ideal MHD equations
%% 
%% When we set the initial magnetic field strength to zero $\magdensv = 0$ no
%% magnetic dynamics come into play and the equations reduces further to the
%% compressible Euler equations.

In astrophysics, the \emph{interstellar medium} (ISM) is the matter that exists
in the space between the star systems in a galaxy. It contains gas in ionic,
atomic, and molecular form, as well as dust and cosmic rays. ISM fills
interstellar space and blends smoothly into the surrounding intergalactic
space. The medium is composed primarily of atomic hydrogen followed by helium
with traces of carbon, oxygen, and nitrogen. Magnetic fields and turbulent
motions also provide pressure in the ISM, and are typically more important
dynamically than the thermal pressure.

Many theoretical models for ISM are basesd on the ideal
\emph{Magneto-Hydrodynamic equations} which are a blend of the \emph{compressible
Euler equations} and of the \emph{Maxwell equations}, describing the hydrodynamics
and (electro-)magnetodynamics, respectively.

In this thesis we solely focus on the hydrodynamics.

\subsubsection{Compressible Euler Equations}

%% euler quation primitive form

%% \begin{align}
%% \label{eqn:comp-euler}
%% \partial_t\dens + \nabla \cdot (\dens\,\velv)   &=  0 \\
%% \dens\, \partial_t \velv + \dens(\velv \cdot \nabla) \velv + \nabla \pres &= \vec{F} \\
%% \partial_t \ener + \nabla \cdot (\ener\,\velv - ...)   &=  0,
%% \end{align}

In 1757 \vip{Leonhard Euler} (1707-1783) published a set of equations for
inviscid flow, known as the Euler equations. They are hyperbolic
conservation equations which model perfect fluids without any interaction of their
constituents.  Hence, we assume no
heat conduction ($\SET^{i0} = \SET^{0i} = 0$), no viscosity
($\SET^{ij} = p\,\mathbb{I}$, $\mu_d = 0$) and no gravity $g = 0$. 
Within the comoving frame the
\emph{stress-energy tensor} $\SET$ reads:
\begin{equation}
\SET^{\alpha\beta} = \text{diag}(\dens\sndsped^2,\pres,\pres,\pres) = \left(\dens + \frac{\pres}{\sndsped^2}\right) \vels^\alpha\vels^\beta + p\,\mathbb{G}^{\alpha\beta}.
\end{equation}
In flat spacetime the \emph{metric tensor} is set to $\mathbb{G} = \text{diag}(-1,1,1,1)$. The total energy
and the number of particles are conserved.
\begin{align}
\partial_\nu \SET^{\mu\nu}  &= 0\\
\partial_\mu (n\,\vels^\mu) &= 0
\end{align}
Taking the non-relativistic limit, we arrive at the conservative form of the
Euler equations.
\begin{align}
\label{eqn:comp-euler-conservative}
\partial_t\dens + \nabla \cdot (\dens\,\velv)   &=  0 \;\; &\text{\emph{mass conservation}}\\[0.4cm]
\partial_t(\dens\,\velv) + \nabla\cdot(\dens\velv\velv^T) + \nabla \pres &= \vec{F}  \;\;&\text{\emph{momentum conservation}}\\[0.4cm]
\partial_t \Ener + \nabla \cdot (\velv\,(\ener + \pres)) &=  0,\;\;&\text{\emph{energy conservation}}
\end{align}
where the total energy $\Ener$ is composed of the internal energy $\Eint$ and
the kinetic energy $\Ekin$.
\begin{equation}
\Ener = \Eint + \Ekin = \frac{\pres}{\gamma -1} + \frac{\dens}{2}\vels^2,
\end{equation}
with $\adiabconst$ being the \emph{adiabatic constant}.

The source term $\vec{F}$ (F for forcing) allows us to perpetually inject a
force field which gets important in the discussion of driven turbulence later
on. See \sec{turbuforcing}.

% \paragraph{Primitive and Consverative Variables}

\paragraph{Equation of State}
If not stated otherwise all simulations follow the \emph{ideal gas law}.
\begin{equation}
\pres = \frac{\sndsped^2}{\gamma} \, \dens = R\,T\,\dens = \frac{R}{c_v} \Eint = (\gamma - 1) \Eint,
\end{equation}
where $R$ is the specific ideal gas constant, $T$ is the gas temperature and $c_v
= \frac{\gamma - 1}{R}$ is the specific heat capacity at constant volume. 

The  $\adiabconst$ is set to
\begin{equation}
\label{eqn:adiabconst}
    \adiabconst = \frac{\heatcapP}{\heatcapV} := \frac{5}{3},
\end{equation}
which represents a mono-atomic gas without interacting forces.

The speed of sound $\sndsped$ is a direct consequence of the ideal gas
equation.  \begin{equation} c^2 = \gamma \frac{\pres}{\dens} := C_P = \text{const.} \end{equation}

During the numerical simulation the equation of state is enforced via the
\emph{polytropic process} (also called \emph{polytropic cooling}) at
every timestep.
\begin{equation}
\pres = C_P \, \dens^\Gamma,
\end{equation}
where the \emph{polytropic exponent} is set to $\Gamma := 1$ which is
equivalent to an isothermal process. A thorough derivation can be found in
\cite[p. 2 ff.]{horedt2004polytropes}.

% \paragraph{Prandtl Number}
% \begin{equation}
% \label{eqn:prandtl}
%     \prandtl = \frac{\shearvisc \; \heatcapP}{\heatconduct}
% \end{equation}

\paragraph{Dimensionless Euler Equations} We want to show that the Euler
equations are invariant to changes of units. This discussion is useful since
most numerical frameworks do not support physical units and rescaled phyiscal
quantities avoid truncation errors due to the limits of floating point
operations. For this, we choose a characteristic length $l_r$, a characteristic
velocity $\vels_r$ and a characterstic density $\dens_r$. Multiplying suitable 
combinations of these constants with the Euler equations yields
\begin{align}
\left[\partial_t\dens + \nabla \cdot (\dens\,\velv)\right] \cdot \frac{l_r}{\dens_r\,\vels_r} &=  0\\[0.4cm]
\left[\partial_t(\dens\,\velv) + \nabla\cdot(\dens\velv\velv^T) + \nabla \pres - \vec{F}\right] \cdot \frac{l_r}{\dens_r\,\vels_r^2} &= 0\\[0.4cm]
\left[\partial_t \Ener + \nabla \cdot (\velv\,(\ener + \pres))\right] \cdot \frac{l_r}{\dens_r\,\vels_r^3} &=  0
\end{align}
We simplify and get
\begin{align}
\partial_{\tilde{t}} \tilde{\dens} + \tilde{\nabla} \cdot (\tilde{\dens}\,\tilde{\velv})   &=  0\\[0.4cm]
\partial_{\tilde{t}}(\tilde{\dens} \,\tilde{\velv}) +      \tilde{\nabla}\cdot(\tilde{\dens}\tilde{\velv}\tilde{\velv}^T) + \tilde{\nabla} \tilde{\pres} - \tilde{\vec{F}}&= 0\\[0.4cm]
\partial_{\tilde{t}} \tilde{\Ener} + \tilde{\nabla} \cdot (\tilde{\velv}\,(\tilde{\ener} + \tilde{\pres})) &=  0,
\end{align}
where $t_r = \frac{l_r}{\vels_r}$ (characteristic time) and
\begin{equation}
\tilde{t} = \frac{t}{t_r},\ 
\tilde{\dens} = \frac{\dens}{\dens_r},\
\tilde{\velv} = \frac{\velv}{\vels_r},\ 
\tilde{\nabla} = l_r\,\nabla,\ 
\tilde{\Ener} = \frac{\Ener}{\dens_r\,\vels_r^2},\ 
\tilde{\pres} = \frac{\pres}{\dens_r\,\vels_r^2},\ 
\tilde{\vec{F}} = \vec{F} \, \frac{l_r}{\dens_r\,\vels_r^2}.
\end{equation}
Conseqently, the dimensionless Euler equations do not change under unit transformation. If not
stated otherwise we drop the tilde sign ($\tilde{\cdot}$) and assume always
dimensionless quantites from now on.

\paragraph{Choice of parameters} One consequence of dimensionless units is the
free choice of parameters. We want to use this feature in choose a sensible set
of parameters.  Considering the Euler equations in conservative form,
\eqn{comp-euler-conservative}, their functions of space and time
\begin{equation}
\dens = \dens(t,x,y,z), \;\;(\dens\velv) = (\dens\velv)(t,x,y,z),\;\;\Ener = \Ener(t,x,y,z)
\end{equation}
are completed with
\begin{equation}
\label{eqn:parameter-choice-completion}
\gamma := 5/3,\;\;R := 1, \langle\dens\rangle := 1,\langle\sndsped\rangle := 1,
\end{equation}
From that we derive
\begin{equation}
C_P = \frac{\sndsped_0^2}{\gamma} = 3/5 = 0.6,\;\;
\langle\pres\rangle = C_P\cdot\langle\dens\rangle = 0.6,\;\;
\langle\Ener\rangle = \frac{\langle\pres\rangle}{\gamma -1} = 0.9,\;\;
\langle\temp\rangle = \frac{\avg{\sndsped}^2}{\gamma\,R} = 0.6,
\end{equation}
where $\avg{\cdot}$ is the \emph{volume-weighted average} or \emph{mean value}
over the domain $\Omega$
\begin{equation}
\label{eqn:domain-average}
    \avg{q} = \frac{\int_\Omega\,q\,d\Omega}{\int_\Omega\,d\Omega}
\end{equation}

If not state otherwise, these set of parameters define the global state at all times.

\subsubsection{Weak Formulation}
A natural way to define a generalized solution of the Euler equations that does
not require differentiability is, to go back to the integral form of the
conservation law,

The basic idea is to take the PDE, multiply it by a smooth \emph{test function},
integrate one or more times over some domain, and then use integration by parts
to move derivatives off the function q and onto the smooth test function. The
result is an equation involving fewer derivatives on q, and hence requiring
less smoothness.

In this section we want to derive the \emph{weak formulation} of the governing
equations. This establishes the basis for the polynomial formulation which is
the core idea of all DG methods.  First, the Euler equations get split up into
terms resembling the independent one temporal and three spatial
dimensions with respect to the linear differential operator.
\begin{equation}
\partial_t \vec{U} + \partial_x\vec{F}(\vec{U})+ \partial_y\vec{G}(\vec{U}) + \partial_z\vec{H}(\vec{U}) = \vec{S},
\end{equation}
where
\begin{align}
    \vec{U} &= (\dens, \dens\vels_1, \dens\vels_2, \dens\vels_3, \Ener)^T\br
    \vec{F}(\vec{U}) &= (\dens\vels_1, \dens\vels_1^2 + \pres, \dens\vels_1\vels_2, \dens\vels_1\vels_3, \vels_1(\Ener + \pres))^T\br
    \vec{G}(\vec{U}) &= (\dens\vels_2, \dens\vels_2\vels_1, \dens\vels_2^2 + \pres, \dens\vels_2\vels_3, \vels_2(\Ener + \pres))^T\br
    \vec{H}(\vec{U}) &= (\dens\vels_3, \dens\vels_3\vels_1, \dens\vels_3\vels_2, \dens\vels_3^2 + \pres, \vels_3(\Ener + \pres))^T\br
    \vec{S} &= (0,f_1,f_2,f_z,0)^T
\end{align}
Defining a vector-valued test function $\vec{\phi} = (0,..,0,\phi_i,0,...,0)^T$ ($i\in{1,...,5}$), multiplying component-wise
with above equation and integrating over the domain $\Omega$ we get
\begin{equation}
    \int_\Omega \left( \partial_tU_i\,\phi^i 
        + \partial_x F_i(\vec{U})\,\phi^i 
        + \partial_y G_i(\vec{U})\,\phi^i 
        + \partial_z H_i(\vec{U})\,\phi^i + S_i \phi^i \right) d^3x = 0
\end{equation}
Integration-by-parts rearranges the integral into a \emph{source term},
\emph{surface term} and \emph{volume term}.
\begin{align}
\label{eqn:weak-formulation}
    \int_\Omega \partial_t U_i\,\phi^i d^3x = \;\;\;\;\;\;\;&\int_\Omega S_i \phi^i d^3x \br
    +& \int_{\partial\Omega} \left(F_i(\vec{U})\,\phi^i\,n_x + G_i(\vec{U})\,\phi^i\,n_z + H_i(\vec{U})\,\phi^i\,n_z \right) d^2x,\nonumber \br
    -& \int_\Omega \left(F_i(\vec{U})\,\partial_x \phi^i + G_i(\vec{U})\,\partial_y \phi^i + H_i(\vec{U})\,\partial_z \phi^i \right) d^3x
\end{align}
where $\vec{n} = (n_x,n_y,n_z)^T$ is the outward surface normal to $\partial\Omega$.

Unfortunately, weak solutions are not unique, and so an additional problem is
to identify which weak solution is the physically correct vanishing-viscosity
solution. Again, one would like to avoid working with the viscous equation
directly, but it turns out that there are other conditions one can impose on
weak solutions that are easier to check and will also pick out the correct
solution. These are called \emph{entropy conditions} by analogy with the gas
dynamics case, where a discontinuity is physically realistic only if the
entropy of the gas increases as it crosses the shock.
