\subsubsection{Turbulent Forcing}
\label{sec-theory:turbulent-forcing}

In equation ... a \emph{source} respectively \emph{forcing} term was introduced
into the Euler equations. Perpetually, at each time step a varying force field
$\vec{F}(t,x,y,z)$ in time and space injects kinetic energy over the whole
physical domain. Inducing natural looking turbulence with the desired
properties is not a trivial task. One commonly used method is by exploiting the
intermittent behaviour of random walks, in particular the
\textsc{Ornstein-Uhlenbeck process}. Since we deploy pseudo-random numbers
the forcing is replicable.

Based on \cite{schmidt2009} we formulate
\begin{align}
  \hat{d\mathbf{f}}(\mathbf{k},t) &= \frac{3}{\sqrt{1-2\zeta+3\zeta^2\,}}\;\left[-\hat{\mathbf{f}}(\mathbf{k},t)\frac{dt}{T} + F_0 \left(\frac{2\sigma^2(\mathbf{k})}{T}\right)^{1/2}\mathbf{P}_\zeta(\mathbf{k})\cdot d\mathbf{W}_t\right]\\
  (P_{ij})(\mathbf{k}) &= \zeta\,P_{ij}^\perp(\mathbf{k}) + (1-\zeta) P_{ij}^\parallel =\zeta\,\delta_{ij} + (1-2\,\zeta) \frac{k_i k_j}{k^2}\\
\end{align}
This acceleration field in fourier space allows us to precisely specify at
which spatial scales we want to apply the forcing as well as the ratio of
compressive and solenoidal modes.

The projection parameter $\zeta \in [0,1]$ sets the relative contribution of
compressible and solenoidal injection rates. If not stated otherwise $\zeta$ is
set to $0.5$. Turning time $T$ and base forcing $F_0$ depend on the general
simulation setup.

The goal is a fully developed turbulence where the turbulent system has reached dynamical equillibrium between
energy inflow, supplied by forcing and outflow caused by small-scale dissipation and shock capturing.
