\section*{Acknowledgment}

First and foremost I want to express my debt of gratitude and deep respect to
my supervisor, Prof. Dr. Stefanie Walch-Gassner, for inviting me to her
workgroup and for her excellent guidance.  Her invaluable advice, far-reaching
insights and open mind were much appreciated.  She granted me a lot of freedom
to pursue my own interests and always had a friendly ear for any thought or
idea. Even when they were stupid, retrospectively.

Likewise I want to thank Prof. Dr.-Ing. Gregor Gassner who encouraged me to
write this thesis in the first place.

Many thanks to PD Dr. Volker Ossenkopf for co-reviewing my work despite his
busy schedule.

\vspace{0.5cm}

I thank my family for their love, support and unlimited patience.

\vspace{3cm}


\begin{tabbing}
\hspace{1cm}    \= \textbf{Examinee} \hspace{4cm}  \= Johannes Markert \\
                \>                                 \> I. Physikalisches Institut \\
                \>                                 \> Universität zu Köln \\
                \>                                 \> markert@ph1.uni-koeln.de  \\\\

                \> \textbf{First Supervisor}       \> Professor Dr. Stefanie Walch-Gassner \\
                \>                                 \> I. Physikalisches Institut \\
                \>                                 \> Universität zu Köln \\
                \>                                 \> walch@ph1.uni-koeln.de  \\\\

                \> \textbf{Second Supervisor}      \> PD Dr. Volker Ossenkopf \\
                \>                                 \> I. Physikalisches Institut \\
                \>                                 \> Universität zu Köln \\
                \>                                 \> ossk@ph1.uni-koeln.de
\end{tabbing}
